Hypercoercions and a Framework for Equivalence of Cast Calculi

### 2019-10-19 paper.tex
abstract

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
\todoKC{rewrite this abstract based on the comparision table.}
  Gradual Typing is great. Implementing gradually typed with blame
  tracking and space efficiency is tricky. There exist two technique
  to do this: coercion and threesome. Coercion is easy to understand,
  and easy enough to implement, but difficult to reason about
  formally. Threesome is hard to understand, easy to implement, and
  easy to reason about formally. We propose hypercoercion, which is
  easy to understand, as easy to implement as coercion, and easy to
  reason about formally.
\end{abstract}

@@

\todo[inline]{What is the main idea behind the organization of
 hypercoercions? -JS
 \\---\\
 Are you asking why hc has three parts? Or why these subsections 
 in this section? -KC
 \\---\\
 three parts -JS
}

This section presents our first contribution, the definition of hypercoercion. 
The functions that construct hypercoercions are deferred to subsections because 
they depend on blame strategies.
And the functions that apply hypercoercions to values are deferred to 
Section~\ref{sec:hypercoercion-correctness} because 
they depends on a new definition of values, which is part of the framework in
Section~\ref{sec:framework}.

@@

Let $ c $ range over hypercoercions. A hypercoercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ range over heads. A head is either a no-op, or a projection.
Let $ m $ range over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to middles $ m \smile m $ in the natural way.
Let $ t $ range over tails. A tail is either a no-op, an injection, or a 
failure. 
Let $ \ell $ range over no-op and labels. It is useful when we introduce 
composition of hypercoercions.

@@

$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

@@

The correctness proof of \lazyD\ hypercoercion is deferred to 
\secref{sec:LDHC-correct} 
because it relies on the space-efficient CEK abstract machine in 
\secref{sec:CEKS}.

@@

	\caption{Space-efficient CEK machine $ \mathcal{S}(C) $}
	\label{machine-cekcc}
\end{figure}
\figref{machine-cekcc} defines, \effCEK{C}, a space-efficient semantics of cast 
calculus in the style of CEK machines.
This machine is similar to \ineffCEK{}. 
The most important differences are in values and continuations.
In \ineffCEK\, we have one value constructor for all casted value.
In \effCEK{C}, however, we do not have this generic value constructor, 
instead we push those casts into the value constructors most of the time and 
use $ \hcvOOinj{P}{v} $ to construct values casted to $ \TOOdyn $.
Non-casted values in \ineffCEK\ correspond to values in \effCEK{C} where the 
casts are identity (i.e. constructed by $ id(T) $).
In \ineffCEK\, a continuation have zero or more casts at the top.
In \effCEK{C}, however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond to 
continuations in \effCEK{C}\ whose casts are identity.
And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $ 
seq(c,c) $.

Let $ v $ range over value. A value is either the unit value, function, pair, 
left injection of sum, right injection of sum, and injection to $ \TOOdyn $.
Cast results ($ r $) and machine states ($ s $) are as before. 
Let $ \kappa $ range over continuations and let $ k $ range over 
pre-continuations. 
A continuation is now a pre-continuation prefixed with a cast.
Pre-continuation are like the continuations in \ineffCEK.

@@

\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	For all instance of Cast ADT $ C $, 
	if $ C $ is \lazyD{} and is a monoid, 
	$ s_{1}, s_{1}' \in \text{\effCEK{C}} $,
	$ s_{2}, s_{2}' \in \text{\ineffCEKD} $,
	$ s_{1} \approx s_{2} $,
	$ \judgeSreduce{C}{s_{1}}{s_{1}'} $, and
	$ \judgeDreduce{s_{2}}{s_{2}'} $, then
%	then there exist $ s_2'' \in \text{\ineffCEKD} $ such that
	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	\begin{enumerate}
%		\item If $ \judgeDreduce{s_1}{s_3} $,
%		then
%		there exist $ s_5 \in \text{\ineffCEKD} $, $ s_4 \in \mathcal{S}(C) $,
%		such that \begin{itemize}
%			\item $ s_5 \approx s_4 $, and
%			\item $ \judgeSreduce{C}{s_2}{s_4} $,
%			\item $ \judgeDreduceTrans{s_3}{s_5} $
%		\end{itemize}
%		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
%		then
%		there exist $ s_3 \in \mathcal{C} $,
%		such that \begin{itemize}
%			\item $ s_3 \approx s_4 $
%			\item $ \judgeDreduceTrans{s_1}{s_3} $
%		\end{itemize}
%	\end{enumerate}

@@

\begin{lemma}[Strong Bi-simulation among $ \mathcal{S}(\cdot) $]
	\label{thm:CEKS-bisim}
	If 
	$ C_1 $ and $ C_2 $ are instances of Lazy D Cast ADT,
	$ s_1, s_1' \in S(C_1) $ and $ s_2,s_2' \in S(C_2) $,
	$ s_1 \approx s_2 $,
	$ \judgeSreduce{C_1}{s_1}{s_1'} $, and
	$ \judgeSreduce{C_2}{s_2}{s_2'} $, then 
	$ s_1' \approx s_2' $

>>>>

%% Abstract
%% Note: \begin{abstract}...\end{abstract} environment must come
%% before \maketitle command
\begin{abstract}
	Designing a space-efficient cast representation that is good for both 
	mechanized metatheory and implementation is challenging. Existing solution 
	solutions have various strengths and weaknesses. We presents a new cast 
	representation, named hypercoercions, that is good for both. On our way to 
	prove the correctness of hypercoercions, we also step toward a framework 
	for proving correctness of cast representations.
\end{abstract}

@@

\todo[inline]{What is the main idea behind the organization of
 hypercoercions? -JS
 \\---\\
 Are you asking why hc has three parts? Or why these subsections 
 in this section? -KC
 \\---\\
 three parts -JS
 \\---\\
 Fixed -KC
}

This section presents our first contribution, the definition of hypercoercion. 
The design of hypercoercion is motivated by observations on coercion normal 
forms: a normal coercion has at most three parts; projections are always at the 
beginning; injections and failures are always at the end. Hypercoercions have 
similar shape: a hypercion has either one part or three parts; for three-part 
hypercoercions, the first part is either a projection or a no-op, while the 
last part is either an injection, a failure, or a no-op.

@@

Let $ c $ range over hypercoercions. A hypercoercion is either 
an identity cast between the dynamic types, or a complex including a head, a 
middle, and a tail. 
Let $ h $ range over heads. A head is either a no-op, or a projection.
Let $ m $ range over middles. There is a one-to-one 
correspondence between middles and type constructors. 
We generalize shallow-consistency to middles $ m \smile m $ in the natural way.
Let $ t $ range over tails. A tail is either a no-op, an injection, or a 
failure.

We define hypercoercion constructs in the following subsections.
But functions that apply hypercoercions to values are deferred to 
Section~\ref{sec:hypercoercion-correctness} because 
they depends on a new definition of values, which is part of the framework in
Section~\ref{sec:framework}.

@@

Let $ \ell $ range over $ \epsilon $ and labels.
$ c_1 \fatsemi^\ell c_2 = c' $ composes two hypercoercions. The target of $ 
c_1 $ and the source of $ c_2 $ might be different, in which case $ \ell $ must 
be a label.
$ m_1 \fatsemi^\ell (m_2,t) = (m',t') $ composes a middle with a 
pair of a middle and a tail. Similarly, the target of $ m_1 $ and the source of 
$ m_2 $ might be different, in which case $ \ell $ must be a label.
$ seq(c_1,c_2) $ composes two coercions where the target type of $ c_1 $ is 
equal to the source type of $ T_2 $.
$ id(T) $ constructs an identity coercion of $ T $ with the help of $ id(P) $.
$ cast(T_1,l,T_2) $ constructs a coercion from a source type, a label, and a 
target type.

@@

@@

	\caption{Space-efficient CEK machine $ \mathcal{S}(C) $}
	\label{machine-cekcc}
\end{figure}
\figref{machine-cekcc} defines, \effCEK{C}, a space-efficient semantics as a 
CEK machine.
This machine is similar to \ineffCEK{}. 
The most important differences are in values and continuations.

Let $ v $ range over value. 
In \ineffCEK\, we have one value constructor for all casted value.
In \effCEK{C}, however, we do not have this generic value constructor, 
instead we push those casts into the value constructors most of the time and 
use $ \hcvOOinj{P}{v} $ to construct values casted to $ \TOOdyn $.
Non-casted values in \ineffCEK\ correspond to values in \effCEK{C} where the 
casts are identity (i.e. constructed by $ id(T) $).

Cast results ($ r $) and machine states ($ s $) are as before.

Let $ \kappa $ range over continuations and let $ k $ range over 
pre-continuations. 
A continuation is now a pre-continuation prefixed with a cast.
Pre-continuation are like the continuations in \ineffCEK.
In \ineffCEK\, a continuation have zero or more casts at the top.
In \effCEK{C}, however, every continuation has exactly one cast at the top.
Continuations in \ineffCEK\ that have no casts at the top correspond to 
continuations in \effCEK{C}\ whose casts are identity.
And continuations in \ineffCEK\ that have many casts at the top correspond to 
continuations in \effCEK{C}\ where those casts are composed by $ 
seq(c,c) $.

@@

\todo[inline]{Standardize theorem statement -KC}
\begin{lemma}[Weak Bisimulation between \effCEK{C} and \ineffCEKD]
	\label{thm:surely-monoidic-reduce}
	For all instance of Cast ADT $ C $, 
	if $ C $ is \lazyD{} and is a monoid, 
	$ s_{1}, s_{1}' \in \text{\effCEK{C}} $,
	$ s_{2}, s_{2}' \in \text{\ineffCEKD} $,
	$ s_{1} \approx s_{2} $,
	$ \judgeSreduce{C}{s_{1}}{s_{1}'} $, and
	$ \judgeDreduce{s_{2}}{s_{2}'} $, then
%	then there exist $ s_2'' \in \text{\ineffCEKD} $ such that
	\[ \exists s_2'' \in \text{\ineffCEKD}: \;
	\judgeDreduceTrans{s_{2}'}{s_{2}''} \;\text{and}\; s_1' \approx s_2'' \]
%	\begin{enumerate}
%		\item If $ \judgeDreduce{s_1}{s_3} $,
%		then
%		there exist $ s_5 \in \text{\ineffCEKD} $, $ s_4 \in \mathcal{S}(C) $,
%		such that \begin{itemize}
%			\item $ s_5 \approx s_4 $, and
%			\item $ \judgeSreduce{C}{s_2}{s_4} $,
%			\item $ \judgeDreduceTrans{s_3}{s_5} $
%		\end{itemize}
%		\item If $ \judgeSreduce{C}{s_2}{s_4} $,
%		then
%		there exist $ s_3 \in \mathcal{C} $,
%		such that \begin{itemize}
%			\item $ s_3 \approx s_4 $
%			\item $ \judgeDreduceTrans{s_1}{s_3} $
%		\end{itemize}
%	\end{enumerate}

@@

\todo[inline]{Standardize theorem statement -KC}
\begin{lemma}[Strong Bi-simulation among $ \mathcal{S}(\cdot) $]
	\label{thm:CEKS-bisim}
	If 
	$ C_1 $ and $ C_2 $ are instances of Lazy D Cast ADT,
	$ s_1, s_1' \in S(C_1) $ and $ s_2,s_2' \in S(C_2) $,
	$ s_1 \approx s_2 $,
	$ \judgeSreduce{C_1}{s_1}{s_1'} $, and
	$ \judgeSreduce{C_2}{s_2}{s_2'} $, then 
	$ s_1' \approx s_2' $
